---
title: "ESM 232 - HW4 Dynamic Models"
author: "Anna Calle and Jamie Miller"
date: "May 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries 
library(tidyverse)
library(deSolve)
library(sensitivity)
library(pse)
```

Consider the following model of forest growth (where forest size in measured in units of carbon (C))

dC/dt = 0 when air temperatures are below 0
dC/dt  = r*C for forests where carbon is below a canopy closure threshold 
dC/dt = g for forests where carbon is above the threshold canopy closure
dC/dt = 0 once carrying capacity is reached.


The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon 
You could think of the canopy closure threshold as the size of the forest at which growth rates change from exponential to linear
You can think of r, as early exponential growth rate and g and the linear growth rate 

#### 1. Implement this model in R (as a differential equation)

#### 2. Run the model for 300 years (using the ODE solver)  starting with an initial forest size of 10 kg/C, and using the following parameters
-canopy closure threshold of 100 kgC 
-a carrying capacity of 500 kgC
-a early growth rate of 0.01 kg/year
-a later growth rate of 2 kg/year
-an air temperature of 20C

Graph the results

```{r}
source("../R/d_forest_growth.R")
initial_C = 10
years = seq(from=1, to=300, by=5)

# create parameter list
newparms = list(canopy_ct = 100, K = 500, r = 0.01, g = 2, air_T = 20)

# run the solver
result = ode(y=initial_C, times=years, func=dforest_growth,parms= newparms)

# look at results
head(result)
colnames(result)=c("year","Cnew")
result_graph <- ggplot(as.data.frame(result), aes(year,Cnew)) +
  geom_point() +
  labs(y="Forest Size (C)", x = "Years")

result_graph

# When does forest size reach carrying capactiy?
# The 'years' sequence was increased from 300 to 500 to check model function. At year 430 the modelled forest size will hit carrying capacity.

```


#### 3. Run a sobol sensitivity analysis that explores how the estimated maximum forest size varies with early and later growth rates. 

Graph the results of the sensitivity analysis as a box plot and a plot of the two sobol indices (S and T)

```{r}
# want to learn about sensitivity to early growth rate (r) and later growth rate (g)
# set the number of parameters
np=10000
r = rnorm(mean=0.01, sd=.2*.01, n=np)
g = rnorm(mean=2, sd=.2*.01, n=np)
X1 = cbind.data.frame(r=r, g=g)

# repeat to get our second set of samples
r = rnorm(mean=0.01, sd=.2*.01, n=np)
g = rnorm(mean=2, sd=.2*.01, n=np)
X2 = cbind.data.frame(r=r, g=g)

# create our sobel object and get sets ofparameters for running the model
sens_C = sobol2007(model = NULL,X1, X2, nboot = 300)

# our parameter sets are
head(sens_C$X)

# turn computing our metrics into a function
compute_metrics = function(result) {
  maxC = max(result$Cnew)
idx = which.max(result$Cnew)
maxyear = result$year[idx]
return(list(maxC=maxC, maxyear=maxyear))}

# turn it into a data frame
result = as.data.frame(result)

# try it on our first parameter set
compute_metrics(result)

# define a wrapper function to do everything we need
c_wrapper = function(r, g, initial_C, years, func) {
    parms = list(canopy_ct = 100, K = 500, r = r, g = g, air_T = 20)
    result = ode(y=initial_C, times= years, func=dforest_growth, parms=parms) 
    colnames(result)=c("year","Cnew")
  # get metrics
    result = as.data.frame(result)
  metrics=compute_metrics(result)
  return(metrics)
}

# use mapply
allresults = mapply(FUN=c_wrapper, r=sens_C$X$r, g=sens_C$X$g,
 MoreArgs=list(initial_C=initial_C, years=years, func=dforest_growth))

# convert to dataframe and change column names
allres = as.data.frame(matrix(unlist(allresults), ncol=2, byrow=T))
colnames(allres) = c("maxC","maxyear")

# create boxplots
tmp = allres %>% gather(key="metric", value="value")
ggplot(tmp, aes(metric, value, col=metric))+geom_boxplot()

```

Compute the sobol indices for each metric

```{r}
# use tell function
sens_C_maxC = sensitivity::tell(sens_C, allres$maxC)

# first-order indicex (main effect without co-variance)
sens_C_maxC$S

# total sensitivity index
sens_C_maxC$T

# plot indices
plot(sens_C_maxC)

# make a data frame for plotting
both = cbind.data.frame(sens_C_maxC$X, allres$maxC, response=sens_C_maxC$y)

# look at response of Max C to the two growth rates
ggplot(both, aes(allres$maxC,r, col=g))+geom_point()+labs(y="Exponential Growth Rate (r)", x="MaxC")
ggplot(both, aes(allres$maxC,r, col=response))+geom_point()+labs(y="Exponential Growth Rate (r)", x="MaxC")

ggplot(both, aes(allres$maxC,g, col=r))+geom_point()+labs(y="Linear Growth Rate (g)", x="MaxC")
ggplot(both, aes(allres$maxC,g, col=response))+geom_point()+labs(y="Linear Growth Rate (g)", x="MaxC")

```



